{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-08-27 ms.tex
Replace deleted material, revert from a3b9a6d
Accidentally removed in a3b9a6d0f10d1b673c0d06679a9c7117169bd577

DS Falster was supported by an Australian Research Council (ARC)
discovery grant (DP110102086). RG FitzJohn was supported by the
Science and Industry Endowment Fund (RP04-174). M Westoby was
supported by a fellowship from the ARC. DSF thanks {\AA}
Br{\"a}nnstr{\"o}m and U Dieckmann for introducing him to the \textsc{ebt}
framework and the eco-evolutionary framework implemented in {\plant}, and
for discussions around prior software. We also thank J Camac, A De Roos,
R Duursma, J Johansson and C Prentice for helpful discussions. The authors 
have no conflicts of interest to declare.
\end{document}

>>>>

DS Falster was supported by an Australian Research Council (ARC)
discovery grant (DP110102086). RG FitzJohn was supported by the
Science and Industry Endowment Fund (RP04-174). M Westoby was
supported by a fellowship from the ARC. DSF thanks {\AA}
Br{\"a}nnstr{\"o}m and U Dieckmann for introducing him to the \textsc{ebt}
framework and the eco-evolutionary framework implemented in {\plant}, and
for discussions around prior software. We also thank J Camac, A De Roos,
R Duursma, J Johansson and C Prentice for helpful discussions. The authors 
have no conflicts of interest to declare.

\clearpage
\bibliography{refs}

\clearpage

\section{Figures}\label{figures}

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/schematic}

\caption{Processes modelled within {\plant} include physiological, population and
evolutionary dynamics.
Dynamics across all three levels are driven by the
physiological sub-model determining an individual's physiological and
demographic rates on the basis of its traits, size and light environment (a).
(b) Competitive hierarchies  are modelled by tracking the height distribution of
individuals within a ``patch''. The intensity
of shading indicates the density of individuals at a given height for
different species (distinguished by colours). A size- and patch-structured
metapopulation consists of a distribution of patches linked by seed dispersal.
Disturbances occasionally remove all vegetation within a patch, resetting the
patch. (c) The traits of the resident species determine the shading environment
across the metapopulation, which in turn determines the fitness landscape.
Resident traits adjust through directional selection
and through the introduction of new species. Adapted from
\citet{Falster-2011}, \citet{Falster-2015}. }

\label{fig:schematic}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/plant.pdf}
\caption{Physiological dynamics for individual plants varying in size,
  trait and light environment. Solid lines are for the high
  \textsc{lma} (leaf mass per unit area) strategy, dashed lines are
  for the low \textsc{lma} strategy.  (a) Growth trajectories are
  influenced by both light environment and traits.  (b) Over time, the
  fraction of living tissue switches from leaf towards sapwood, with
  high \textsc{lma} species having relatively more mass in leaf than
  low \textsc{lma} species. (c) Size-dependent growth rates
  (derivative of panel (a)) peak at around 5m high, but this position
  varies with both trait and light level.  (d) Declines in growth rate
  with light vary with plant size and traits.  Whole plant light
  compensation points (indicated by dots) emerge as the point where
  growth rate is zero (x intercept).  The light level used in panels
  (a) and (c) is indicated with coloured ticks.  See Appendix
  \ref{sec:plant} for further details and code.}
\label{fig:plant}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/patch.pdf}
\caption{Population dynamics for two species competing within a patch.
  Red and blue lines are low and high \textsc{lma} species from
  Fig. \ref{fig:plant}.  Lines in panel (a) show growth trajectories
  for individuals on the boundary of successive ``cohorts'' for each
  species, using a schedule of introduction times generated by
  {\plant}'s adaptive algorithm.  The initial growth rate advantage of
  the low \textsc{lma} species (see Figure \ref{fig:plant}c) means it
  quickly over tops the high \textsc{lma} species, suppressing its
  growth.  Self-thinning creates space for the high \textsc{lma}
  species to establish below the canopy of the faster species. Panel
  (b) shows the same trajectories as in panel (a), but with shading
  indicating the density $N$ of individuals at given size and patch
  age (light = low density, dark = high density).  (c) The leaf area
  index at ground level for each species.  The dotted line is the
  total leaf area index (summed across both species), and the lines on
  the right side are average when integrated over the whole
  metapopulation structure.  See Appendix \ref{sec:patch} for further
  details and code.}
\label{fig:patch}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/emergent.pdf}
\caption{Emergent size-distributions and demography across a metapopulation.
Grey lines show patterns within individual patches differing in time
since disturbance (age), with blue lines indicating patches of a
specific age. Red lines show average relationship across the metapopulation, 
obtained by weighting the grey lines by patch and individual abundance. 
Panels show (a) patterns of abundance, (b) leaf area and
(c) height growth rate against size. Units of panels (a) and
(b) are amount per unit height per unit ground area.  
See Appendix \ref{sec:emergent} for further details and code.}
\label{fig:emergent}
\end{figure}

\newpage

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/fitness.pdf}
\caption{Example fitness landscape for \textsc{lma} (leaf mass per
  unit leaf area), showing potential for stable coexistence of
  multiple types.  Fitness is the log of the population growth rate --
  at zero species are at equilibrium, above zero they will increase in
  density.  (a) Landscape generated by a single species (indicated by
  dots).  At \textsc{lma} of 0.1 (the dashed line, open dot) there is
  directional selection for smaller \textsc{lma}.  At \textsc{lma} of
  0.08 (solid line, filled dot) there is a branching point (see inset
  for more detail).  (b) Landscape generated by two species, holding
  the first at \textsc{lma} of 0.08.  Introducing a species at the
  maximum fitness value from panel (a) reduces the region of positive
  fitness considerably (dashed line, open dot).  Dotted lines show
  subsequent invasions and replacements of the second species, the
  solid line shows a stable fitness landscape where both species exist
  on a peak in the fitness landscape.  At this point no further
  invasion is possible.  See Appendix \ref{sec:fitness} for further
  details and code.}
\label{fig:fitness}
\end{figure}

\clearpage
\setcounter{secnumdepth}{1}

\begin{appendices}\label{sec:appendices}

\section{Plant physiological model}\label{sec:FFW16}

See attached file \url{vignettes/physiology.pdf}

\section{Modelling demography of plants, patches and metapopulations}\label{sec:demography}

See attached file \url{vignettes/demography.pdf}

\section{Plant level properties}\label{sec:plant}

\emph{Note}: due to Manuscript Central not accepting \textsc{html}
files, we have had to convert this (and the following appendices) to
\textsc{pdf}s.  \textsc{html} versions are available on the wiki:
https://github.com/traitecoevo/plant\_paper/wiki --- the \textsc{html}
versions have less pixelated graphics and are easier to copy and
paste out of.

See attached file \url{vignettes/plant.pdf}, 
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/plant.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/plant}

\section{Cohort spacing algorithm}\label{sec:cohort-spacing}

See attached file \url{vignettes/cohort_spacing.pdf},
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/cohort_spacing.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/cohort_spacing}

\section{Finding demographic equilibrium}\label{sec:equilibrium}

See attached file \url{vignettes/equilibrium.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/equilibrium.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/equilibrium}

\section{Patch level dynamics}\label{sec:patch}

See attached file \url{vignettes/patch.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/patch.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/patch}

\section{Patch level emergent properties}\label{sec:emergent}

See attached file \url{vignettes/emergent.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/emergent.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/emergent}

\section{Calculating fitness}\label{sec:fitness}

See attached file \url{vignettes/fitness.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/fitness.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/fitness}

\section{Modifying parameters of the physiological model}\label{sec:parameters}
See attached file \url{vignettes/parameters.pdf}
source:
\smurl{https://github.com/traitecoevo/plant\_paper/blob/master/vignettes/src/parameters.R}
html: \smurl{https://github.com/traitecoevo/plant\_paper/wiki/parameters}

\end{appendices}
\end{document}
